---
title: "p8105_hw1_mss2284"
author: "Maya Spaur"
date: "9/13/2019"
output: github_document
---

## Problem 1

```{r Problem 1 code chunk 1}
library(tidyverse)

problem1_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = norm_samp > 0,
  vec_char = character(length=8),
  vec_factor = factor(c("1", "2", "3", "3", "2", "1", "1", "2")),
)

mean(pull(problem1_df, norm_samp))
mean(pull(problem1_df, vec_logical))
mean(pull(problem1_df, vec_char))
mean(pull(problem1_df, vec_factor))
```

The means were able to be calculated for norm_samp and vec_logical, but not for vec_char and vec_factor.

Converting the variables to numeric:

```{r Problem 1 code chunk 2, results='hide', warning=FALSE}

as.numeric(pull(problem1_df, vec_logical))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))

as.numeric(pull(problem1_df, vec_logical))*(pull(problem1_df, norm_samp))

as.factor(pull(problem1_df, vec_logical))*(pull(problem1_df, norm_samp))

as.numeric(as.factor(pull(problem1_df, vec_logical)))*(pull(problem1_df, norm_samp))

```
As numeric was able to convert the logical variable (assigns 0 to false and 1 to TRUE). It didn't work for the character variable, it could not be converted to a numeric variable. The factor variable was able to be converted to 3 levels.

## Problem 2


```{r Problem 2 code chunk 1}

problem2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical_2 = x + y > 1,
  as.numeric(vec_logical_2),
  as.factor(vec_logical_2),
)

```

The dataset has a number of rows`r nrow(problem2_df)`, and number of columns `r ncol(problem2_df)`.

The mean of the variable x is `r mean(pull(problem2_df, x))`. The median of the variable x is `r median(pull(problem2_df, x))`. The standard deviation of the variable x is `r sd(pull(problem2_df, x))`.


The proportion of cases for which x + y is greater than one is `r mean(pull(problem2_df, vec_logical_2))`.


Scatterplot

```{r scatter}

ggplot(problem2_df, aes(x = x, y = y, color = vec_logical_2)) + geom_point()

ggsave

ggplot(problem2_df, aes(x = x, y = y, color = as.numeric(problem2_df$vec_logical_2))) + geom_point()

ggplot(problem2_df, aes(x = x, y = y, color = as.factor(problem2_df$vec_logical_2))) + geom_point()

```
 
 For the scatterplot with the as.numeric variable, the color scale is represented as a gradient from light to dark blue to reflect the continuous variable. For the plot using the logical variable and factor variable, there were two colors, to represent the binary (logical) variable, blue for true and orange for false.